### 2012-09-19 process.tex
initial latex files with 2 reference papers
rbac-slr.tex

\section{Methodology and Process} \label{sec:process}

Reference can be Andrew Meneely's Validating Software Metrics: A Spectrum of Philosophies paper.