### 2012-09-27 process.tex
Introduction and RQs

\section{Methodology and Process} \label{sec:process}

We collect papers on extended role-based access control~\cite{ferraiolo:rbac}.
We select 27 papers out of XXX papers.

\subsection{Research Questions} \label{subsec:research questions}
We have following research questions.
\begin{itemize}
	\item RQ1. What are problems in current RBAC model to propose extended RBAC models?
	\item RQ2. What are reasons to propose RBAC extensions? 
	\item RQ3. Which extended features of RBAC model are proposed?
	\item RQ4. How they provide evidence to show that their model work in practice? For example, they provide a prototype to run a real example.
	\item RQ5. What is criteria for evaluation? We investigate how they evaluate their research work.

\end{itemize}

For RQ1-RQ3, we compare of proposed extended models.
For RQ4, we compare quality of completeness of the proposed models.
Beyond modeling of extended RBAC models, an prototype shows that their models work in practice and improves the quality
of the completeness of research papers. For RQ5, we compare criteria for evaluation of extended RBAC models.

\subsection{Categorization} \label{subsec:categorization}

This section describes categorization of papers based on specific extended features as follows. 
\begin{itemize}
	\item Temporal-related-constraints~\cite{mossakowski03:temporal,aich07:STARBAC,kumar06:strbac,samuel07:spatio-temporal,ray07:spatio,chandran05:llt,aich09:role,bertino01:trbac, li08:fine, joshi05:generalized, chen08:spatio-temporal}
	\item Inheritance~\cite{ren10:machine}
	\item Location~\cite{chandran05:llt}
	\item Delegation~\cite{hasebe10:capability}
	\item General constraints~\cite{alam06:constraint,yao08:task}
	\item Context~\cite{tzelepi01:flexible,haibo05:context,cholewka00:acontext-sensitive,huang06:pervasive,motta03:contextual,bao08:role}						
	\item General extension~\cite{han08:extended,zhang06:collaborative}
	\item Combination with other access control models such as Task-based access control~\cite{yao08:task,oh03:task,zhou07:network, oh00:task} and Agent-based access control~\cite{yamazaki04:designing}
\end{itemize}