### 2013-02-03 paper/rbac-slr.tex
Re-factor of location of latex elements.  Now main paper is stored under paper and within that each major section of the paper is stored under sections directory.

\documentclass[preprint,12pt]{elsarticle}

\usepackage{amssymb}

\journal{TBD}

%% DOCUMENT %%
\begin{document}

%% 
\begin{frontmatter}

\title{A Systematic Literature Review on Role-Based Access Control}

\author{}

\address{}

\input{sections/abstract}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
\end{keyword}

\end{frontmatter}

%% SECTIONS %%
\input{sections/intro}
\input{sections/process}

%% REFERENCES %%
\bibliographystyle{elsarticle-num}
\bibliography{sections/rbacslr}

\end{document}