### 2013-02-06 paper/sections/terms.tex
Adding terms section.

\section{Core Role Based Access Control} \label{sec:core-rbac}

The NIST RBAC model is proposed by Ferraiolo et al.. The model
consists of four basic components:

\begin{itemize}
	\item a set of users \emph{Users}: A user can be a person or an agent.
	\item  a set of roles \emph{Roles}: A role is a collection of permissions to perform a specific job function in an organization.
	\item a set of permissions \emph{Permissions}: A permission refers to an access mode that
can be exercised on an object in the system and a session
relates a user to possibly many roles.
	\item a set of sessions \emph{Sessions}: In each session, a
user can be assigned to some of the roles, only when the corresponding role is enabled for activation for that time.		
\end{itemize}

In the RBAC, a user can exercise a permission only if the user are assigned to a role.
In addition to the four basic components, two functions are defined:
the user role assignment (UA) and the role
permission assignment (PA) functions.
UA models assignment of users to roles.
PA models assignment of permissions to roles.

%Each user incorporates a session    
%The user function maps each session to a
%single user, whereas the role function establishes a
%mapping between a session and a set of roles activated
%by the corresponding user in the session.

\section{Terms and Definitions} \label{sec:terms}

We found that different definitions for the same terms. Therefore, we next describe
terms and definitions.

\begin{itemize}
%	\item Access Control Model:
%	\item Access Control Policy:
%	\item Access Control Requirements:
%	\item Rules:
%	\item Attributes-Based Access Control:
	\item Task-Based Access Control:
%	\item Purpose-Based Access Control:
	\item Agent-Based Access Control:						
\end{itemize}

\begin{itemize}

%	\item Role Hierarchies:

	\item Obligations: obligations are requirements, which should be fulfilled before or after authorization decision is enforced.
	Consider that a user has permission to access specific resources. For example, obligation is that the user should complete
	her/his office duty before accessing the resources.

	\item Inheritance:
	Inheritance defines an inheritance relationship among attributes such as roles. For example, the role structure for a company use
	employee role for employees. Department manager may inherit all permissions of the employee role. (Role-Based Access Control by F. Ferrailolo et al.)

	\item Static Separation of Duty (SSoD): SSoD restricts the conflicting-role assignments statically that are associated with a user. On situations
	where multiple roles can be associated with a single user and roles $Role_A$ and $Role_B$ are conflicting each other, no permission is given to a user who is assigned to both $Role_A$ and $Role_B$. SSoD is known to be too rigid for practical use in cases where a user should have permissions as either $Role_A$ and $Role_B$. (Role-Based Access Control by F. Ferrailolo et al.)

	\item Dynamic Separation of Duty: Dynamic SoD (DSD) is known to be
more flexible than SSD. DSD restricts the conflicting-role assignments dynamically that are associated with a user. On situations
	where multiple roles can be associated with a single user and , given a context, roles $Role_A$ and $Role_B$ are conflicting each other dynamically, no permission is given to a user. (Role-Based Access Control by F. Ferrailolo et al.)
	\end{itemize}

We define temporal and spatial constraints as follows:

\begin{itemize}
	\item Temporal Constraints: Temporal constraints are time-based constraints in specifying access
	control policies. For example, in organizations, 	periodic temporal durations are enforced while a
	specific role is permitted to conduct an action. Consider that part-time employee works only from 9:00 a.m. to 3:00 p.m.
	In such cases, the part-time employee role should access required resources during the interval. 
	Temporal constraints can incorporate either on roles, user-role assignments, or role-permission assignments.   
	(Role-Based Access Control by F. Ferrailolo et al.)

	\item Spatial Constraints: Spatial constraints are location-based constraints in specifying access
	control policies. For example, in organizations, 	locations are enforced while a
	specific role is permitted to conduct an action. Consider that part-time employee works only in specific location.
	In such cases, the part-time employee role should access required resources only when the user is in the location. 
	Spatial constraints can incorporate either on roles, user-role assignments, or role-permission assignments.

\end{itemize}