### 2013-02-09 paper/sections/conclusion.tex
Adding some updated writing and notes on each section and what to put.

\section{Conclusion} \label{sec:conclusion}

P1: Mention core RBAC, note specific results for each research question

P2: Mention bigger results that cut across and any other peritentn information