### 2013-02-09 paper/sections/results.tex

\section{Results} \label{sec:results}

P1: Discuss focus on going through each research question individually, and combining results across at the end of the process

P2: List the research questions

S1: RQ1 and discuss by category and general results

S2: RQ2 and discuss by category and general results

S3: RQ3 and discuss by category and general results

S4: RQ4 and discuss by category and general results

S5: RQ5 and discuss by category and general results