### 2013-02-19 paper/rbac-slr.tex
Adds outline for tracking. Adds some preliminary research question sections and analysis. Adds more to results section.

\documentclass[3p,12pt]{elsarticle}

\usepackage{amssymb}
\usepackage{hyperref}

\journal{TBD}

%% DOCUMENT %%
\begin{document}

%% 
\begin{frontmatter}

\title{A Systematic Literature Review on Role-Based Access Control}

\author{}

\address{}

\input{sections/abstract}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
\end{keyword}

\end{frontmatter}

\section{Outline}

\begin{enumerate}
\item Introduction
    \begin{itemize}
  	\item RBAC history, RBAC as a standard
  	\item Audience that should care about our review
  	\item Brief explanation of what has led to extensions cropping up
  	\item Paper organization
    \end{itemize}
\item Background
    \begin{itemize}
   	\item What is RBAC
   	\item Core RBAC and it's entities, and what it tries to solve
    \end{itemize}
\item Process
\item Results
    \begin{itemize}
  	\item Data used in process, results of process for paper count
  	\item Categorization that came out applying process (based on word usage)
  	\item Results for each research question (raw data)
    \end{itemize}
\item Analysis
    \begin{itemize}
  	\item Explanation and definitions of categories based on results
  	\item Examination of results in relation to each research question and larger trends drawn from data
  	\item Examination of cross research question concerns an
    \end{itemize}
\item Conclusion
\end{enumerate}

%% SECTIONS %%
\input{sections/intro}
\input{sections/terms}
\input{sections/process}
\input{sections/results}
\input{sections/analysis}

%% REFERENCES %%
\bibliographystyle{elsarticle-num}
\bibliography{sections/rbacslr}

\end{document}