### 2013-02-27 paper/rbac-slr.tex
Updates to paper list into appendix. Adds structured abstract. Moves paper counts into process.

\documentclass[3p,12pt]{elsarticle}

\usepackage{amssymb}
\usepackage{hyperref}

\journal{TBD}

%% DOCUMENT %%
\begin{document}

%% 
\begin{frontmatter}

\title{A Systematic Literature Review on Extensions to Role Based Access Control}

\author{Eric D. Helms, JeeHyun Hwang, Tao Xie, Laurie Williams}
\address{North Carolina State University}
\address{Department of Computer Science}
\address{890 Oval Drive, Box 8206}
\address{Raleigh, NC 27695-2858}

\input{sections/abstract}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
RBAC \sep access control \sep systematic literature review
\end{keyword}

\end{frontmatter}

%% SECTIONS %%
\input{sections/intro}
\input{sections/core_rbac}
\input{sections/process}
%%\input{sections/results}
\input{sections/analysis}
\input{sections/conclusion}

%% REFERENCES %%
\bibliographystyle{elsarticle-num}
\bibliography{sections/rbacslr}

%% Outline - REMOVE ME %%
\section{Outline}

\begin{enumerate}
\item Introduction
    \begin{itemize}
  	\item RBAC history, RBAC as a standard
  	\item Audience that should care about our review
  	\item Brief explanation of what has led to extensions cropping up
  	\item Paper organization
    \end{itemize}
\item Background
    \begin{itemize}
   	\item What is RBAC
   	\item Core RBAC and it's entities, and what it tries to solve
    \end{itemize}
\item Process
\item Results
    \begin{itemize}
  	\item Data used in process, results of process for paper count
  	\item Categorization that came out applying process (based on word usage)
  	\item Results for each research question (raw data)
    \end{itemize}
\item Analysis
    \begin{itemize}
  	\item Explanation and definitions of categories based on results
  	\item Examination of results in relation to each research question and larger trends drawn from data
  	\item Examination of cross research question concerns an
    \end{itemize}
\item Conclusion
\end{enumerate}

\input{sections/appendix}

\end{document}