### 2013-02-27 paper/sections/appendix.tex

\section{Appendix} \label{sec:appendix}

\begin{table}
\centering
\begin{tabular}{|p{12.5cm}|p{3cm}|}
\hline
\textbf{Paper} & \textbf{Category}
\\\hline
A constraint based role based access control in the SECTET a model-driven approach, 2006 \cite{alam06:constraint} & Constraint \\\hline

A flexible content and context-based access control model for multimedia medical image database systems, 2001 \cite{tzelepi01:flexible} & Context \\\hline

A context-sensitive access control model and prototype implementation, 2000 \cite{cholewka00:acontext-sensitive} & Context \\\hline

A Context, Rule and Role-Based Access Control Model In Enterprise Pervasive Computing Environment, 2006 \cite{huang06:pervasive} & Context \\\hline

A contextual role-based access control authorization model for electronic patient record Motta, 2003 \cite{motta03:contextual} & Context \\\hline

A Role and Context Based Access Control Model with UML, 2008 \cite{bao08:role} & Context \\\hline

An extended RBAC model based on granular logic, 2008 \cite{jian2008extended} & Context \\\hline

Designing an agent-based RBAC system for dynamic security policy, 2004 \cite{yamazaki104:designing} & Context \\\hline

An extended RBAC model based on granular logic, 2008 \cite{han08:extended} & Context \\\hline

Leveraging Access Control Mechanism of Android Smartphone Using Context-Related Role-Based Access Control Model, 2011 \cite{thein2011leveraging} & Context \\\hline

CRBAC: Imposing multi-grained constraints on the RBAC model in the multi-application environment, 2009 \cite{zou2009crbac} & Context \\\hline

ROBAC: Scalable Role and Organization Based Access Control Models, 2006 \cite{zhang06:collaborative} & Organizational \\\hline

Privacy-aware role-based access control, 2007 \cite{ni2010privacy} & Privacy \\\hline

PuRBAC: Purpose-Aware Role-Based Access Control, 2008 \cite{masoumzadeh2008purbac} & Privacy \\\hline

A Flexible Role- and Resource-Based Access Control Model, 2008 \cite{zhao2008flexible} & Resource \\\hline

GEO-RBAC: a spatially aware RBAC, 2005 \cite{damian2007geo} & Spatial \\\hline

LRBAC: A location-aware role-based access control model, 2006 \cite{ray07:spatio} & Spatial \\\hline

Spatial role-based access control model for wireless networks, 2003 \cite{hansen2003spatial} & Spatial \\\hline

STARBAC: Spatiotemporal role based access control, 2007 \cite{aich07:STARBAC} & Spatio-Temporal \\\hline

On spatio-temporal constraints and inheritance in role-based access control, 2008 \cite{chen08:spatio-temporal} & Spatio-Temporal \\\hline

A framework for specification and verification of generalized spatio-temporal role based access control model, 2007 \cite{samuel07:spatio-temporal} & Spatio-Temporal \\\hline

LoT-RBAC: A Location and Time-Based RBAC Model, 2005 \cite{chandran05:llt} & Spatio-Temporal \\\hline

A Spatio-temporal Role-Based Access Control Model, 2007 \cite{ray07:spatio} & Spatio-Temporal \\\hline

Role Based Access Control with Spatiotemporal Context for Mobile Applications, 2009 \cite{aich09:role} & Spatio-Temporal \\\hline

A Task-Role Based Access Control Model with Multi-Constraints, 2008 \cite{yao2008task} & Task \\\hline

Team and Task Based RBAC Access Control Model, 2009 \cite{zhou2007team} & Task \\\hline

Task-role-based access control model, 2003 \cite{oh2003task} & Task \\\hline

A generalized temporal role-based access control model, 2005 \cite{joshi05:generalized} & Temporal \\\hline

\end{tabular}
\caption{Primary sources grouped by categorization}
\label{tab:categorization}
\end{table}