### 2013-03-03 paper/sections/appendix.tex
Adds heirarchy table in appendix. Adds structured abstract.

\section{Appendix} \label{sec:appendix}

\begin{sidewaystable}
\centering
\vline
\begin{tabularx}{\linewidth}{ | X | X | X | X | }
\cline{1-4}

\multicolumn{1}{ c }{\multirow{2}{*} {\rotatebox{90}{ Context } }} & 
\multicolumn{3}{ |l| }{A constraint based role based access control in the SECTET a model-driven approach, 2006 \cite{alam06:constraint}} \\ \cline{2-4}

\multicolumn{1}{ c }{}  &
\multicolumn{3}{ |l| }{A flexible content and context-based access control model for multimedia medical image database systems, 2001 \cite{tzelepi01:flexible}} \\ \cline{2-4}

\multicolumn{1}{ c }{}  &
\multicolumn{3}{ |l| }{A constraint based role based access control in the SECTET a model-driven approach, 2006 \cite{alam06:constraint}} \\ \cline{2-4}

\multicolumn{1}{ c }{}  &
\multicolumn{3}{ |l| }{A flexible content and context-based access control model for multimedia medical image database systems, 2001 \cite{tzelepi01:flexible}} \\

\multicolumn{1}{ c }{}  &
\multicolumn{3}{ |l| }{A context-sensitive access control model and prototype implementation, 2000 \cite{cholewka00:acontext-sensitive}} \\ \cline{2-4}

\multicolumn{1}{ c }{}  &
\multicolumn{3}{ |l| }{A Context, Rule and Role-Based Access Control Model In Enterprise Pervasive Computing Environment, 2006 \cite{huang06:pervasive}} \\ \cline{2-4}

\multicolumn{1}{ c }{}  &
\multicolumn{3}{ |l| }{A contextual role-based access control authorization model for electronic patient record Motta, 2003 \cite{motta03:contextual}} \\ \cline{2-4}

\multicolumn{1}{ c }{}  &
\multicolumn{3}{ |l| }{A Role and Context Based Access Control Model with UML, 2008 \cite{bao08:role}} \\ \cline{2-4}

\multicolumn{1}{ c }{}  &
\multicolumn{3}{ |l| }{An extended RBAC model based on granular logic, 2008 \cite{jian2008extended}} \\ \cline{2-4}

\multicolumn{1}{ c }{}  &
\multicolumn{3}{ |l| }{Designing an agent-based RBAC system for dynamic security policy, 2004 \cite{yamazaki104:designing}} \\ \cline{2-4}

\multicolumn{1}{ c }{}  &
\multicolumn{3}{ |l| }{An extended RBAC model based on granular logic, 2008 \cite{han08:extended}} \\ \cline{2-4}

\multicolumn{1}{ c }{}  &
\multicolumn{3}{ |l| }{Leveraging Access Control Mechanism of Android Smartphone Using Context-Related Role-Based Access Control Model, 2011 \cite{thein2011leveraging}} \\ \cline{2-4}

\multicolumn{1}{ c }{}  &
\multicolumn{3}{ |l| }{CRBAC: Imposing multi-grained constraints on the RBAC model in the multi-application environment, 2009 \cite{zou2009crbac}} \\ \cline{2-4}

\multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{{\multirow{1}{*}{\rotatebox{90}{ Org. }}}} &
\multicolumn{2}{ |l| }{ROBAC: Scalable Role and Organization Based Access Control Models, 2006 \cite{zhang06:collaborative}} \\[2.75ex]  \cline{2-4}

\multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{{\multirow{2}{*}{\rotatebox{90}{ Priv. }}}} &
\multicolumn{2}{ |l| }{Privacy-aware role-based access control, 2007 \cite{ni2010privacy}} \\[1ex] \cline{3-4}

\multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
\multicolumn{2}{ |l| }{PuRBAC: Purpose-Aware Role-Based Access Control, 2008 \cite{masoumzadeh2008purbac}} \\[1ex] \cline{2-4}

\multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{{\multirow{1}{*}{\rotatebox{90}{ Res. }}}} &
\multicolumn{2}{ |l| }{A Flexible Role- and Resource-Based Access Control Model, 2008 \cite{zhao2008flexible}} \\[2.5ex]  \cline{2-4}

\multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{{\multirow{3}{*}{\rotatebox{90}{ Task }}}} &
\multicolumn{2}{ |l| }{A Task-Role Based Access Control Model with Multi-Constraints, 2008 \cite{yao2008task}} \\ \cline{3-4}

\multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
\multicolumn{2}{ |l| }{Team and Task Based RBAC Access Control Model, 2009 \cite{zhou2007team}} \\ \cline{3-4}

\multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
\multicolumn{2}{ |l| }{Task-role-based access control model, 2003 \cite{oh2003task}} \\ \cline{2-4}

\multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{{\multirow{10}{*}{\rotatebox{90}{ Spatio-Temporal }}}} &
\multicolumn{2}{ |l| }{STARBAC: Spatiotemporal role based access control, 2007 \cite{aich07:STARBAC}} \\[0.25ex] \cline{3-4}

\multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
\multicolumn{2}{ |l| }{On spatio-temporal constraints and inheritance in role-based access control, 2008 \cite{chen08:spatio-temporal}} \\[0.25ex] \cline{3-4}

\multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
\multicolumn{2}{ |l| }{A framework for specification and verification of generalized spatio-temporal role based access control model, 2007 \cite{samuel07:spatio-temporal}} \\[0.25ex] \cline{3-4}

\multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
\multicolumn{2}{ |l| }{LoT-RBAC: A Location and Time-Based RBAC Model, 2005 \cite{chandran05:llt}} \\[0.25ex] \cline{3-4}

\multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
\multicolumn{2}{ |l| }{A Spatio-temporal Role-Based Access Control Model, 2007 \cite{ray07:spatio}} \\[0.25ex] \cline{3-4}

\multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
\multicolumn{2}{ |l| }{Role Based Access Control with Spatiotemporal Context for Mobile Applications, 2009 \cite{aich09:role}} \\[0.25ex] \cline{3-4}

\multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{} & \multicolumn{1}{ |c }{{\multirow{3}{*}{\rotatebox{90}{ Spatial }}}} &
\multicolumn{1}{ |l| }{GEO-RBAC: a spatially aware RBAC, 2005 \cite{damian2007geo}} \\[0.25ex] \cline{4-4}

\multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{} & \multicolumn{1}{ |c }{}  &
\multicolumn{1}{ |l| }{LRBAC: A location-aware role-based access control model, 2006 \cite{ray07:spatio}} \\[0.25ex] \cline{4-4}

\multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{} & \multicolumn{1}{ |c }{}  &
\multicolumn{1}{ |l| }{Spatial role-based access control model for wireless networks, 2003 \cite{hansen2003spatial}} \\[0.25ex] \cline{3-4}

\multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{} & \multicolumn{1}{ |c }{{\multirow{3}{*}{\rotatebox{90}{ Temp. }}}} &
\multicolumn{1}{ |l| }{A generalized temporal role-based access control model, 2005 \cite{joshi05:generalized}} \\[5ex]

\cline{1-4}
\end{tabularx}
\vline
\caption{Primary sources grouped by categorization}
\label{tab:categorization}
\end{sidewaystable}