### 2013-03-24 paper/rbac-slr.tex
Breaks each research quesiton into it's own major section.

\documentclass[3p,12pt]{elsarticle}

\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{rotating}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage{comment}
\usepackage{array}
\usepackage{epstopdf}
\usepackage{enumitem}

\journal{TBD}

%% DOCUMENT %%
\begin{document}

%% 
\begin{frontmatter}

\title{A Systematic Literature Review on Extensions to Role Based Access Control}

\author{Eric D. Helms, JeeHyun Hwang, Tao Xie, Laurie Williams}
\address{North Carolina State University}
\address{Department of Computer Science}
\address{890 Oval Drive, Box 8206}
\address{Raleigh, NC 27695-2858}

\input{sections/abstract}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
RBAC \sep access control \sep systematic literature review
\end{keyword}

\end{frontmatter}

%% SECTIONS %%
\input{sections/intro}
\input{sections/core_rbac}
\input{sections/process}
\input{sections/categorization}
\input{sections/analysis}
\input{sections/conclusion}

%% REFERENCES %%
\bibliographystyle{elsarticle-num}
\bibliography{sections/rbacslr}

\input{sections/appendix}

\end{document}