### 2014-02-09 paper/sections/appendix.tex

\section{Appendix} \label{sec:appendix}

\begin{table}
\centering
\caption{Primary sources by classification and properties (Implementation Type / Domain / Evaluation Type(s))}
\vline
\begin{tabularx}{\textwidth}{ | c | c | c | c | l | l | }
\cline{1-4}

\multicolumn{1}{ c }{\multirow{2}{*} {\rotatebox{90}{ Context } }} & 
\multicolumn{3}{ | p{.9\textwidth} | }{\cite{tzelepi01:flexible} A flexible content and context-based access control model for multimedia medical image database systems, 2001 (None / Medical / Scenario)} \\ \cline{2-4}

\multicolumn{1}{ c }{}  &
\multicolumn{3}{ | p{.95\textwidth} | }{\cite{haibo05:context} A Context-Aware Role-Based Access Control Model for Web Services, 2005 (None / Web / None)} \\ \cline{2-4}

\multicolumn{1}{ c }{}  &
\multicolumn{3}{ | p{.95\textwidth} | }{\cite{cholewka00:acontext-sensitive} A context-sensitive access control model and prototype implementation, 2000 (Prototype / Enterprise Workflow / Scenarios) } \\ \cline{2-4}

\multicolumn{1}{ c }{}  &
\multicolumn{3}{ | p{.95\textwidth} | }{\cite{huang06:pervasive} A Context, Rule and Role-Based Access Control Model In Enterprise Pervasive Computing Environment, 2006 (Prototype / Pervasive Computing / Scenarios) } \\ \cline{2-4}

\multicolumn{1}{ c }{}  &
\multicolumn{3}{ | p{.95\textwidth} | }{\cite{motta03:contextual} A contextual role-based access control authorization model for electronic patient record Motta, 2003 (Enterprise / Medical / Case Study) } \\ \cline{2-4}

\multicolumn{1}{ c }{}  &
\multicolumn{3}{ | p{.95\textwidth} | }{\cite{bao08:role} A Role and Context Based Access Control Model with UML, 2008 (Prototype / Enterprise Workflow / Complexity, Comparison to Standard, Scenarios)} \\ \cline{2-4}

\multicolumn{1}{ c }{}  &
\multicolumn{3}{ | p{.95\textwidth} | }{\cite{jian2008extended} An extended RBAC model based on granular logic, 2008 (None / Organizations / Scenarios)} \\ \cline{2-4}

\multicolumn{1}{ c }{}  &
\multicolumn{3}{ | p{.95\textwidth} | }{\cite{yamazaki04:designing} Designing an agent-based RBAC system for dynamic security policy, 2004 (None / Organizations / Scenarios)} \\ \cline{2-4}

\multicolumn{1}{ c }{}  &
\multicolumn{3}{ | p{.95\textwidth} | }{\cite{thein2011leveraging} Leveraging Access Control Mechanism of Android Smartphone Using Context-Related Role-Based Access Control Model, 2011 (None / Mobile Computing / Scenarios)} \\ \cline{2-4}

\multicolumn{1}{ c }{}  &
\multicolumn{3}{ | p{.95\textwidth} | }{\cite{zou2009crbac} CRBAC: Imposing multi-grained constraints on the RBAC model in the multi-application environment, 2009 (None / Mobile Computing / None)} \\ \cline{2-4}

\multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{{\multirow{1}{*}{\rotatebox{90}{Const. }}}} &
\multicolumn{2}{ | p{.7\textwidth} | }{\cite{alam06:constraint} A constraint based role based access control in the SECTET a model-driven approach, 2006 } \\[4ex] \cline{2-4}

\multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{{\multirow{1}{*}{\rotatebox{90}{ Org. }}}} &
\multicolumn{2}{ | p{.7\textwidth} | }{\cite{zhang06:collaborative} ROBAC: Scalable Role and Organization Based Access Control Models, 2006 } \\[2.75ex]  \cline{2-4}

\multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{{\multirow{2}{*}{\rotatebox{90}{Priv. }}}} &
\multicolumn{2}{ | p{.7\textwidth} | }{\cite{ni2010privacy} Privacy-aware role-based access control, 2007 } \\[1ex] \cline{3-4}

\multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
\multicolumn{2}{ | p{.7\textwidth} | }{\cite{masoumzadeh2008purbac} PuRBAC: Purpose-Aware Role-Based Access Control, 2008 } \\[1ex] \cline{2-4}

\multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{{\multirow{3}{*}{\rotatebox{90}{ Task }}}} &
\multicolumn{2}{ | p{.7\textwidth} | }{\cite{yao2008task} A Task-Role Based Access Control Model with Multi-Constraints, 2008 } \\ \cline{3-4}

\multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
\multicolumn{2}{ | p{.7\textwidth} | }{\cite{zhou2007team} Team and Task Based RBAC Access Control Model, 2009 } \\ \cline{3-4}

\multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
\multicolumn{2}{ | p{.7\textwidth} | }{\cite{oh2003task} Task-role-based access control model, 2003 } \\ \cline{2-4}

\multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{{\multirow{10}{*}{\rotatebox{90}{ Spatio-Temporal }}}} &
\multicolumn{2}{ | p{.7\textwidth} | }{\cite{aich07:STARBAC} STARBAC: Spatiotemporal role based access control, 2007 } \\[0.25ex] \cline{3-4}

\multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
\multicolumn{2}{ | p{.7\textwidth} | }{\cite{chen08:spatio-temporal} On spatio-temporal constraints and inheritance in role-based access control, 2008 } \\[0.25ex] \cline{3-4}

\multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
\multicolumn{2}{ | p{.7\textwidth} | }{\cite{samuel07:spatio-temporal} A framework for specification and verification of generalized spatio-temporal role based access control model, 2007 } \\[0.25ex] \cline{3-4}

\multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
\multicolumn{2}{ | p{.7\textwidth} | }{\cite{chandran05:llt} LoT-RBAC: A Location and Time-Based RBAC Model, 2005 } \\[0.25ex] \cline{3-4}

\multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
\multicolumn{2}{ | p{.7\textwidth} | }{\cite{ray07:spatio} A Spatio-temporal Role-Based Access Control Model, 2007 } \\[0.25ex] \cline{3-4}

\multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
\multicolumn{2}{ | p{.7\textwidth} | }{\cite{aich09:role} Role Based Access Control with Spatiotemporal Context for Mobile Applications, 2009 } \\[0.25ex] \cline{3-4}

\multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{} & \multicolumn{1}{ |c }{{\multirow{3}{*}{\rotatebox{90}{ Spatial }}}} &
\multicolumn{1}{ | p{.6\textwidth} | }{\cite{damiani2007geo} GEO-RBAC: a spatially aware RBAC, 2005 } \\[0.25ex] \cline{4-4}

\multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{} & \multicolumn{1}{ |c }{}  &
\multicolumn{1}{ | p{.6\textwidth} | }{\cite{ray2006lrbac} LRBAC: A location-aware role-based access control model, 2006 } \\[0.25ex] \cline{4-4}

\multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{} & \multicolumn{1}{ |c }{}  &
\multicolumn{1}{ | p{.6\textwidth} | }{\cite{hansen2003spatial} Spatial role-based access control model for wireless networks, 2003 } \\[0.25ex] \cline{3-4}

\multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{} & \multicolumn{1}{ |c }{{\multirow{3}{*}{\rotatebox{90}{ Temp. }}}} &
\multicolumn{1}{ | p{.6\textwidth} | }{\cite{joshi05:generalized} A generalized temporal role-based access control model, 2005 } \\[4ex]

\cline{1-4}
\end{tabularx}
\vline
\label{tab:categorization}
\end{table}