### 2014-02-09 paper/sections/conclusion.tex
update category names
change updated categories naming.
-> categories with Const, Org, Priv, Task, Spatio-Temporal, Spatial, and Temporal falling under the general category of context.

\section{Conclusion} \label{sec:conclusion}

The RBAC extension models were revealed to fall into a number of categories with Const, Org, Priv, Task, Spatio-Temporal, Spatial, and Temporal falling under the general category of context. 
The categories each had properties specific to their implementation, but were seen to generalize to being specialized instances of context tailored to the entities or actions the categories covered.
A number of domains were identified as being the motivations behind needing extensions to the RBAC Reference model.  The domains, such as healthcare, presenting new challenges the previous models were not
required to design for.  Our literature review showed that the state of RBAC extension model evaluation needs focused from the research community given most model evaluations seen within the papers were
based on hypothetical situations with little to no case studies or implementations in practice.