### 2014-02-11 paper/sections/analysis.tex
Updating counts in appendix and other data tables.

\section{RQ2: Motivations} \label{sec:motivations}

\textit{What are motivations behind the RBAC extension models?}
\\

During the paper reading phase, we identify motivations of the RBAC extension models by category. We identify motivations by looking for issues regarding why the RBAC reference model was inadequate and how the authors addressed the issues. For example, the papers in the Spatial Category argue that the RBAC reference model does not support spatial constraints, which are used in determining users' access decisions based on spatial dimensions (i.e., location). To address this issue, the authors developed the RBAC extension models by incorporating additional spatial entities and their relations with existing entities of the RBAC Reference Model.

\subsection{Results}

We describe motivations behind the RBAC extension models by category as follows:

\begin{itemize}

  \item \textbf{Context}: The RBAC Reference Model does not support the notion of context constraints  related to changes in environments. Therefore, RBAC belongs to the static access control model, which may not capture changes in environments.

  \item \textbf{Constraint}: The RBAC Reference Model has limitations on its features such as delegation and role hierarchy. For example, partial inheritance in role hierarchy needs to be developed.

  \item \textbf{Organizational}: The RBAC extension model does not handle RBAC administrative tasks efficiently across multiple organizations. The model needs to reduce the administrative complexity of RBAC across multiple organizations.

  \item \textbf{Privacy}: The RBAC extension model does not support the notion of privacy. For example, the model lacks components, constraints, and obligations to handle privacy in RBAC.

  \item \textbf{Task}: The RBAC extension model does not support the notion of a task, team, purpose, and organizational roles, which help specify a business activity in enterprises.

  \item \textbf{Spatio-Temporal}: The RBAC extension model does not support spatial (location-based) and temporal (time-based) constraints, which specify role-assignment, role-activation and permissions based on location and time.

  \item \textbf{Spatial}: The RBAC extension model does not support the notion of spatial (location-based)  constraints, which specify role-assignment and permissions based on location.

  \item \textbf{Temporal}: The RBAC extension model does not support the notion of temporal (time-based)  constraints, which specify role-assignments and permissions based on time.

\end{itemize}

\subsection{Analysis and Discussion}

The RBAC standard provides the RBAC Reference Model, which is an abstraction on top of authorization based on $UA$ and $PA$. RBAC Reference Model presents four entities: roles, sessions, users, and permissions, and their relations. While the RBAC Reference Model is considered fundamental in any RBAC systems, the RBAC Reference Model has limitations on providing features such as dealing with context for emerging applications such as healthcare and mobile devices. For example, in cases where the RBAC extension model does not support the notion of specific constraints, the RBAC extension model incorporates additional entities with regards to context/constraints and their relations with existing entities of the RBAC Reference Model. The RBAC extension model provides general agreement on the definition for specific features. The RBAC extension model helps simplify theoretical modeling and practical implementation of features in the model.

%The Const category claims that the RBAC extension model has limitations on its features. The RBAC extension model in this category incorporates modification of existing features in the RBAC Reference Model. For models in this category, changes occur within the RBAC Reference Model.
%The Task category claims that the RBAC extension model does not support the notion of task, team, purpose, and organizational roles. Because task-based access control has been developed, the RBAC extension model in this category is a combination of the RBAC Reference Model and another access control model, which uses model-specific attributes and their relations. For this combination, new associations between attributes of two models should be introduced. Models in this category require new associations between attributes of different access control models to work together.
%
%
%The Context, Org, Priv, Spatio-Temporal, Spatial, and Temp categories claim that the RBAC extension model does not support the notion of specific constraints. The RBAC extension model in these categories incorporates additional entities with regards to context/constraints and their relations with existing entities of the RBAC Reference Model. The extension model does not change the RBAC Reference Model.
%
%The Const category claims that the RBAC extension model has limitations on its features. The RBAC extension model in this category incorporates modification of existing features in the RBAC Reference Model. For models in this category, changes occur within the RBAC Reference Model.
%
%The Task category claims that the RBAC extension model does not support the notion of task, team, purpose, and organizational roles. Because task-based access control has been developed, the RBAC extension model in this category is a combination of the RBAC Reference Model and another access control model, which uses model-specific attributes and their relations. For this combination, new associations between attributes of two models should be introduced. Models in this category require new associations between attributes of different access control models to work together.

\greybox{Motivations for RBAC extension models vary across category but are rooted in adding new entities or relationships to allow authorization flexibility in meeting the demands of emerging requirements.}

\section{RQ3: Implementations} \label{sec:implementations}

\textit{How are RBAC extension models implemented by their authors?}
\\

When designing and proposing a model targeted at a feature that is rooted in practical
usage by real software systems, implementing the model is evidence that the
proposed model can work in practice. The presence of an actual implementation substantiates the robustness of the design and/or the need for the extension.
We analyzed the primary sources to see how many
proposed models actually had implementations associated with them.
We quantified whether the implementation was enterprise (i.e., an implementation used in a production environment) was for a real system) or a prototype implementation (i.e., proof of concept to demonstrate the feasibility).

\subsection{Results}

Table \ref{tab:implementations} shows the breakdown of implementations found within the primary sources.
Of the 27 papers surveyed, RBAC extension models in four papers were prototypes developed by the authors whereas RBAC extension models in three papers were claimed to be implemented within a real
system. The remaining 20 papers provide no
mention of an implementation.

\begin{table}
\centering
\caption{Implementation types found and the count of primary sources}
\vspace{0.1 in}
\begin{tabular*}{.9\linewidth}{| l | p{.5\linewidth} | r | }
\cline{1-3}

\textbf{Implementation Type} & \textbf{Papers} & \textbf{Count} \\ \cline{1-3}
Enterprise
&
\cite{motta03:contextual}
\cite{aich09:role}
\cite{yao2008task}
&
3 \\ \cline{1-3}
Prototype
& 
\cite{cholewka00:acontext-sensitive}
\cite{huang06:pervasive}
\cite{bao08:role}
\cite{zhou2007team}
&
4 \\ \cline{1-3}
None
&
\cite{alam06:constraint}
\cite{tzelepi01:flexible}
\cite{yamazaki04:designing}
\cite{jian2008extended}
\cite{thein2011leveraging}
\cite{zou2009crbac}
\cite{haibo05:context}
\cite{zhang06:collaborative}
\cite{ni2010privacy}
\cite{masoumzadeh2008purbac}
\cite{damiani2007geo}
\cite{ray2006lrbac}
\cite{hansen2003spatial}
\cite{aich07:STARBAC}
\cite{chen08:spatio-temporal}
\cite{samuel07:spatio-temporal}
\cite{chandran05:llt}
\cite{ray07:spatio}
\cite{oh2003task}
\cite{joshi05:generalized}
&
20 \\

\cline{1-3}
\end{tabular*}
\label{tab:implementations}
\end{table}

\subsection{Analysis and Discussion}

The RBAC standard is designed such that when practitioners implemented RBAC into their systems, the RBAC standard demonstrates reasonable assurance being based off a well thought out model.
As extensions to the RBAC Reference Mode come along, thought and time would be given to how features and nuances of their models may impact implementation to achieve the same goals as the original standard. 
The primary sources showed a lack of implementation
with over 70\% of the models having no notion of attempting to implement them. Prototype implementation shows the feasibility of the RBAC extension models in the four papers. Of the models that produced an implementation within the enterprise world, two were from within the medical domain and one was implemented using web application technologies.

\greybox{Research on RBAC extension models shows a lack of implementations in the real world scenarios that the models are ultimately designed for.}

\section{RQ4: Evaluations} \label{sec:evaluations}

\textit{How are RBAC extension models evaluated by their authors?}
\\

The 27 primary sources were examined for evidence that evaluations of the proposed model were presented by the model authors. We identified a set of evaluation types found within  each of the primary sources and provide a list of the evaluation types and which primary sources provided which type. A primary source may use multiple evaluation types. For example, Aich et al. conducted performance and complexity analyses to show the effectiveness of their proposed RBAC extension model \cite{aich09:role}.

\subsection{Results}

Table \ref{tab:evaluations} shows evaluation types.
Based on the diverse evaluation criteria, 12 models presented no evidence of an evaluation. Fourteen models presented example scenarios
and how application of their model would apply and resolve the situation. Six models provided some form of performance
or complexity analysis of their model.  The performance and complexity analysis included graphs of the model's time to determine calculate authorization 
as the number of entities grew, and the size of the role space for the extension model compared to the RBAC Reference Model. 
Five models provided mathematical descriptions and analysis as a way to provide evaluation in the form of completeness. 
The most widely used evaluation approach was providing sample scenarios with accompanying workflows of how the extension model
would tackle those scenarios. 
%Much is left to the reader to assume of these types of evaluations, as the authors do not explicitly state
%or show how the Reference Model is deficient in tackling said scenarios.

\begin{table}
\centering
\caption{Evaluation types by primary source}
\vspace{0.1 in}
\begin{tabular*}{.9\linewidth}{| p{.45\linewidth} | p{.3\linewidth} | r | }
\cline{1-3}

\textbf{Evaluation Type} & \textbf{Papers} & \textbf{Count} \\ \cline{1-3}
Time-Based Performance
&
\cite{ni2010privacy}
\cite{aich09:role}
&
2 \\ \cline{1-3}
Complexity Analysis
&
\cite{bao08:role}
\cite{zhang06:collaborative}
\cite{chen08:spatio-temporal}
\cite{aich07:STARBAC}
&
4 \\ \cline{1-3}
Comparison to Standard RBAC
& 
\cite{bao08:role}
\cite{zou2009crbac}
\cite{zhang06:collaborative}
\cite{zhao2008flexible}
\cite{ray07:spatio}
&
5 \\ \cline{1-3}
Mathematical Modeling
& 
\cite{damiani2007geo}
\cite{hansen2003spatial}
\cite{aich07:STARBAC}
\cite{chen08:spatio-temporal}
\cite{joshi05:generalized}
&
5 \\ \cline{1-3}
Example Scenarios of the Model in Action
& 
\cite{alam06:constraint}
\cite{tzelepi01:flexible}
\cite{cholewka00:acontext-sensitive}
\cite{huang06:pervasive}
\cite{bao08:role}
\cite{jian2008extended}
\cite{yamazaki04:designing}
\cite{zou2009crbac}
\cite{ray07:spatio}
\cite{samuel07:spatio-temporal}
\cite{joshi05:generalized}
\cite{yao2008task}
\cite{zhou2007team}
\cite{oh2003task}
&
14 \\ \cline{1-3}
Case Study of the Model in Practice
&
\cite{motta03:contextual}
&
1 \\ \cline{1-3}
None
&
\cite{masoumzadeh2008purbac}
\cite{chandran05:llt}
\cite{ray2006lrbac}
&
3 \\ \cline{1-3}

\end{tabular*}
\label{tab:evaluations}
\end{table}

\subsection{Analysis and Discussion}

When proposing an access control model, providing an evaluation of the model is a key component in establishing the validity of the model. 
Further, in the case of extensions to the RBAC Reference Model, the model is accompanied by validation of the model as a stand-alone
access control model and in comparison to the model upon which the enhancements are being made. The results show that robust evaluations of
extension models are lacking.

The primary source of validation a developer or practitioner may encounter is a qualitative discussion of real-world
scenarios and how the proposed model can tackle those situations. In rare cases, five primary sources, the model 
authors provide some discussion of how the RBAC Reference Model is deficient in tackling the scenario.
In Table \ref{tab:evaluations}, we observe that one paper includes a case study to examine how the proposed model works in practice. Discussions
of how an extension model handles a real-world scenario provides developers and practitioners anecdotal evidence at best
for what types of situations the proposed model could handle. Further, by not providing a comparison to the RBAC Reference Model, developers may be left implementing a more complex model to address their requirements when the RBAC Reference Model would
have sufficed.  Further, given the nature of access control models as grounded in application to enterprise implementations, 
case studies of a model in action provide developers with evidence that the model works as intended when applied.

When looking for an enterprise ready access control mechanism, developers must balance usability with security. 
Two of the primary sources examined provided time-based performance analysis of their extension model compared
to the RBAC Reference Model. This inclusion of time analysis provides some assurances to developers that any non-functional
requirements surrounding time to compute authorizations compete or beat the RBAC Reference Model. Further, four of the models provided
some form of complexity of their model. This complexity analysis plays a key role in the management of the access control mechanism
over the course of the models implementation lifetime. As the number of roles, users and additional entities grows, developers will 
need to ensure non-functional requirements are met that deal with the ability for a system administrator to effectively manage these
entities.

\greybox{RBAC extension model research lacks comprehensive evaluation of the models both in theory and in practice.}

\section{RQ5: Domains} \label{sec:domains}

\textit{What domains have been targeted by RBAC extension models?}
\\

Business needs have historically driven RBAC research and development.  The primary mode of evaluation for
model extensions has been the presentation of business scenarios in various domains and how the model
uniquely handles those particular scenarios.  Thus, looking for trends in the domains used in the example
scenarios might serve to illuminate a trend worth further examination into the reason for the explosion of
RBAC extensions.  We identified domains presented within the primary sources by looking for example
scenarios cast within a particular domain or mention of domain requirements within the body of the paper.

\subsection{Results}

Table \ref{tab:domains} shows domains mentioned and their associated sources.

\begin{table}
\centering
\caption{Domains by primary source}
\vspace{0.1 in}
\begin{tabular*}{.9\linewidth}{| p{.45\linewidth} | p{.3\linewidth} | r | }
\cline{1-3}

\textbf{Domain} & \textbf{Papers} & \textbf{Count} \\ \cline{1-3}
Medical Domain
&
\cite{alam06:constraint}
\cite{tzelepi01:flexible}
\cite{motta03:contextual}
\cite{ni2010privacy}
\cite{damiani2007geo}
\cite{hansen2003spatial}
\cite{samuel07:spatio-temporal}
\cite{aich09:role}
\cite{zhou2007team}
&
9 \\ \cline{1-3}
Pervasive Computing Environments
& 
\cite{huang06:pervasive}
\cite{chen08:spatio-temporal}
\cite{ray07:spatio}
&
3 \\ \cline{1-3}
Web Applications
& 
\cite{masoumzadeh2008purbac}
\cite{haibo05:context}
&
1 \\ \cline{1-3}
Mobile Computing
& 
\cite{thein2011leveraging}
\cite{zou2009crbac}
\cite{chandran05:llt}
\cite{aich09:role}
\cite{ray2006lrbac}
&
4 \\ \cline{1-3}
Organizations with Many Sub-departments
& 
\cite{yamazaki04:designing}
\cite{yao2008task}
&
2 \\ \cline{1-3}
Enterprise Workflows
&
\cite{cholewka00:acontext-sensitive}
\cite{bao08:role}
\cite{zhang06:collaborative}
\cite{oh2003task}
\cite{joshi05:generalized}
&
5 \\ \cline{1-3}
None
&
\cite{jian2008extended}
\cite{aich2007:STARBAC}
&
2 \\ \cline{1-3}

\end{tabular*}
\label{tab:domains}
\end{table}

The predominant domain for which extension models have been generated for is that of the medical domain with 9 of 27 mentioning scenarios or requirements of that industry.
Mobile computing and enterprise workflows were each represented by five papers claiming to be influenced by the requirements for access control within these domains. The final set
of domains was pervasive computing environments and large-scale organizations with three each and web applications with one.  Two papers~\cite{jian2008extended, aich07:STARBAC} do not mention domains explicitly since Aich et al. \cite{aich09:role} fall under both the medical domain and mobile computing.

\subsection{Analysis and Discussion}

The medical domain produced the largest selection of papers when analyzing the domains influencing the proposals of RBAC extension models.  
Moreover, we observed that the categories associated with papers identifying the medical domain were not limited to one or two but cut across
each of the eight categories except for the Organizational category. 
The cross-category nature of the medical domain papers appears indicative of the complex nature of medical applications and requirements.
Given the growth of the research and development of medical applications over the past decade this result does not appear to be surprising. However,
the RBAC standard was originally created to reduce cost and increase interoperability - two goals of current regulation around the standardization
of electronic health record systems. The large number of proposed models, and the cross-category result stand in direct opposition of the goals
of both the RBAC standard and current regulations.

The RBAC standard has been re-enforced by the economic impact that standardization has had on enterprises needing to apply access control.  The
inclusion of extension models targeted at the enterprise workflow domain is indicative of the expansion of requirements for enterprises. Developers
and researchers would take care when looking at extension models designed to address the newer requirements of enterprise workflows to
achieve the same economic implementation and maintainability benefits the RBAC Reference Model presents.

Figure~\ref{fig:dist_domains} shows domain distribution by 3-year period.
We observed that medical and enterprise domain papers constantly appear for every period in Figure~\ref{fig:dist_domains}.
For a domain that has roots in medical and enterprise computing, protecting the data of both through access controls is paramount given their ubiquity. 
Mobile computing has seen a dramatic increase in the number of available devices, operating systems and applications since 1997 when the first smart phone was introduced \cite{buck2013new}. Since 2004, the domain analysis results produced five papers that targeted extensions that are designed to address the requirements of mobile computing.

\greybox{RBAC extension models were found targeting domains such as medical, pervasive computing, web applications, enterprise workflows and mobile computing.}

\begin{figure}[ht]
    \centering
        \includegraphics[width=4.0in]{sections/dist_domains_byYear.png}
\vspace{-0.2 in}
    \caption{\label{fig:dist_domains}Domain distribution by 3-year period. Y-axis represents the number of corresponding domain papers.
    X-axis represents 3-year period from 2000 to 2011.}
\end{figure}

\section{RQ6: Generalizations} \label{sec:generalizations}

\textit{What commonalities exist across RBAC extension models?}
\\

During the paper reading phase, we identified commonalities within the primary sources
by looking for formal modeling of extending RBAC extension models.

\subsection{Results}

The RBAC standard is used in various aspects of computer systems. To reduce efforts for modeling access control used in various applications, researchers often focus on developing generalized core concepts of access control.
The authors use propositional logic to describe access control model across all categorizations. Propositional logic is concerned with propositions and their logical relationships. In propositional logic, simple (i.e., atomic) or compound condition at given context is evaluated to true or false based on specified rules and access control logic. Researchers and practitioners are concerned to extend limited set of propositions specific to core RBAC to meet real-world scenarios such as dynamic constraints, temporal, or spatial constraints.

\subsection{Analysis and Discussion}

Since NIST proposed RBAC standard using propositional logic, researchers describe extended RBAC models using propositional logic. Given context, propositional logic is used to evaluate true or false based on specified rules and access control logic. As access control is typically evaluated to either true or false based on predicates, propositional logic is sufficient to describe key ideas and definitions. Propositional logic specifies what combination of attributes values a request must satisfy to be authorized to roles. 
Combination of attribute values may handle constraints such as temporal and spatial constraints. For example, we describe user-role authorization as $ae$ $\Rightarrow$ $r$ using propositional logic where $ae$ is an attribute expression and $r$ is a role. Propositional logic for RBAC has two parts. The left hand side (LHS) is an attribute expression $ae$. The right hand side (RHS) is a role. If a request satisfies the attribute expression, a user $u$ in the request is authorized to the role specified in RHS.

We found that 27 papers of RBAC extensions use propositional logic to describe its extended model.
However, propositional logic has limitations. While this logic is simple, this logic does not support for reasoning about RBAC, which helps
reduce the administrative complexity of associations such as user- role associations.
To support for reasoning of access control, one may describe the RBAC extended models in first-order logic. 
Given an RBAC extension model, Samuel et al.~\cite{samuel07:spatio-temporal} proposed verification of the model using a specification language, which is based first-order logic. This logic is sufficient to model RBAC and extended RBAC for reasoning. Moreover, this logic supports for concise and elegant formulation of the Reference Model and its relation.  First-Order logic is expressive enough to concisely represent access control systems. First-Order logic uses relations, variables, and quantifiers.