### 2014-04-15 paper/reviewer_responses.tex
Adding reviewer responses.

Reviewer #1:

The paper aims to present a systematic literature review (SLR) on RBAC
extensions.  However, a couple of serious issues raise doubts if the
literature review was actually "systematic". In particular, the
process described in the paper is intransparent and hides important
information that is needed to actually assess the paper.

Section 3 ("Methodology and Process") does not describe a systematic
literature review process, as claimed by the authors.  Instead, this
section merely gives an overview of the search and screening part of
the SLR.  However, even this part is roughly sketched only and omits
significant information. For example, it is never explained why
certain search engines resp. certain digital libraries were included
in the review while other very important libraries are ignored. Why
were Springer (SpringerLink) and Elsevier (ScienceDirect), two of the
most important publishers of scientific research, excluded from the
review ?

Moreover, Section 3 does neither reveal the search terms that were
used in the review, nor does it describe the actual search procedure.
Are the 3 terms from Table 1 the final search terms ? If so, these
terms are way too generic. Why didn't the authors focus the search on
their research questions (defined in Section 1) ?

Furthermore, every search engine works differently - how did the
search strings look like ?  Were the search terms applied to paper
titles and abstracts only or did the authors conduct a full-text
search ?  Which type of papers did the authors target (research
papers, opinion papers, white papers, conference papers, journal
papers)? Was the search restricted to a certain publication period ?

In steps 2-4 of the search and screening process (Section 3), the
paper reports on pairs of authors serving as raters on publications
(elimination steps). The paper also refers to (substantial)
negotiation rounds to establish agreement on whether to
include/exclude a publication item. However, the paper does not report
on the levels of agreement/disagreement between the raters over all
items. The elimination quantities from Table 2 are too coarse
grained. However, such IRR analyses are recommended by Kitchenham's
guidelines (allegedly adopted by the authors).

In addition, the stopping criteria are strange . Why was the search
stopped if the result set was greater than 1000 ? How could you
guarantee that the most relevant papers are amongst the first 1000
publications found in the search and not among the papers that were
simply ignored? The paper says "the result set size stopping criteria
was chosen due to a technical limitation of some search
engines". However, the alleged "technical limitations" are never
discussed in more detail. Everybody is well aware of certain "quota"
for searches per hour or searches per day that are enforced by
different scientific search engines. This, however, is certainly no
valid reason for truncating a systematic (!) literature review.

Moreover, stopping "if the last ten results did not contain the search
term phrase within the title" is certainly not the best option: a
potentially large number of relevant publications will not be found.
An alternative would be to establish a quasi-gold standard and search
until a threshold is reached.

Section 3.2.1.: Why was it a requirement that the paper titles must
include certain terms, such as "model"?  I fear that this requirement
is excessively limiting and potentially excluding relevant papers.
Moreover, why is a reference to the term "model" important in the
first place? This is never motivated, no rationale for this
requirement is provided to the reader.

Section 3.2.3.: The use of "conditional inclusion criteria" is not
clear to me (given that they are also not covered by the SLR
guidelines allegedly adopted by the authors). Moreover, the statement
that "without this conditional include, the paper list size" would
have been too "small" indicates that the authors deviated from the
initial procedure (protocol) without justifying or reporting this
deviation.  Arbitrarily changing the inclusion/exclusion criteria to
expand the paper base is not a systematic approach.  With respect to
the content elimination criteria: what does "inspired by real world
example" mean exactly ?

The search strategy results in a relatively small corpus of 27 papers,
searching the papers' references via snowballing could help here.

Moreover, the paper includes a number of inconsistencies that are
never explained. For example, the data from Table 3 and Table 4 are
inconsistent -- in Table 4, papers [21] and [23] are missing while [36]
is newly introduced.

Section 4/RQ1: It is not clear how the authors established the common
terms and how they systematically tagged the included papers in terms
of a content analysis. Also, the wording indicates that all of this
happened during paper reading (ad hoc), rather than as a preparatory
step.

Section 5/RQ2: I do not see how the answers to this RQ are related to
the SLR. The section rather reads like a discussion based on the
authors' intuitions or conjectures.

Section 7/RQ4: Why do the authors only distinguish between
enterprise/prototype "implementations" of RBAC approaches? Wouldn't
it be more interesting to distinguish different "implementation
techniques"?  Moreover, often an implementation alone doesn't say much
about the actual applicability of a certain approach. What about other
evaluation techniques such as case studies, static and dynamic
analysis, formal proofs, or controlled experiments for example ?

Section 9/RQ6: The authors' suggestion of using "propositional
logic" is a pure discussion item and not even remotely backed by the
data extracted from the SLR.

In chapter 10.1. (Discussion) the paper proposes the use of metrics
developed by Hu & Scarfone when choosing an RBAC extension model. Why
this metric in particular? Did the authors perform a prior comparative
analysis of different metrics?

From the results presented in the paper, i have major doubts that the
search terms were adequate and that the search was conducted
thoroughly.  For example, the authors did not find a single
contribution published in the proceedings of the the premier access
control conference -- the ACM Symposium on Access Control Models and
Technologies (SACMAT) which was established in 2001 as a successor of
the ACM RBAC workshop. Moreover, although the paper references some
papers published in ACM TISSEC, many other papers from ACM TISSEC that
do also present extensions to the RBAC reference model were not found
in the SLR.

In its current state, the paper has to be classified as work in
progress. It does neither increase our knowledge of RBAC nor of RBAC
extensions.  The literature review reported in this paper may serve as
a pre-study for an actual SLR. However, the paper needs a lot of work
before it can be accepted for publication in a journal. In particular,
the paper is intransparent and includes a number of serious
methodological flaws.

Minor:

- Section 2 should be shortened to at most one page

- The definitions of static and dynamic separation of duty in Section
2.3 are not correct.

- why are the paper references from Tables 3, 4, and 5 sorted randomly
and not in an ascending order ?

- Table 5 includes a broken literature reference (probably to [29])

- additional figures/visualizations would help to improve the paper

- a number of entries in the References section are incomplete

- the paper includes a number of spelling errors, grammatical errors,
and punctuation flaws.

- Table 6 (Appendix A) should be revised, esp. the structure needs to
be improved

Reviewer #2:

The topic analysed in this paper is interesting and the literature review has been carried out by following a formal approach.

Nevertheles, the authors should justify some decisions made in the systematic review.

Their search strategy should be justified.
        - Why have they selected that search chain?
        - What is the criteria for selecting those sources (Google Shoolar, etc.) and no other?
        - Why were they using as a title elimination criteria that title contains a reference to 'model' and not to "policy", etc. ?

On the other hand, the authors have defined the classification criteria by observation when they were reading the papers. And also, they have defined the comparison items by themself. I consider that these criteria should be defined based on existing frameworks for comparing access control policies and if they do not exist, justify why they propose these criteria.